---
title       : Hypothesis Testing of the Rising of Earth Surface Temperature
subtitle    : 
author      : Bertrand Rigaldies
job         : Data Science Enthusiast
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

```{r global_options, echo=FALSE, warning=FALSE, message=FALSE, error=TRUE}
require(dplyr)
require(lubridate)
require(ggplot2)
require(xtable)
#opts_chunk$set(warning=FALSE, message=FALSE, error=TRUE)
```

## Climate Change - Is your country warming up?

1. Some climate change questions that have captivated us:
    + Have temperatures been rising in your country?
    + Since when, or during what period?
    + Is the temperature rise statistical significant?
1. Data Exploration
1. Statistical Inference
1. Check out my [Shinyapp](https://brigaldies.shinyapps.io/ClimateChange/).

--- .class #id 

## The Data: Earth Surface Temperature Data

```{r load-data, echo=FALSE, results='hide', warning=FALSE, message=FALSE, error=TRUE}
source(file = './server.R')
#data_raw <- read.csv('./GlobalLandTemperaturesByMajorCity.csv')
#data_raw <- na.omit(data_raw) # Remove incomplete records
#data_raw <- data_raw %>% 
#    mutate(Date = ymd(dt)) %>% 
#    mutate(Year = year(Date))
```

```{r sample-data, echo=FALSE, results='hide'}
# We'll use a data sample:
country <- 'United States'
sample_data <- data0 %>% filter(Country == country & City == 'New York')
sample_data_min_year <- min(sample_data$Year)
sample_data_max_year <- max(sample_data$Year)
sample_data <- sample_data %>% 
    filter(Year >= sample_data_max_year) %>% 
    select(dt, Country, City, AverageTemperature, AverageTemperatureUncertainty, Latitude, Longitude)
```

1. Source: [Kaggle](https://www.kaggle.com/berkeleyearth/climate-change-earth-surface-temperature-data)
2. Sample data (New York, United States, `r sample_data_max_year`)

```{r show-sample-data, echo=FALSE, results='asis'}
names(sample_data) <- c("Date", "Country", "City", "AvgTemperature", "AvgTempUncertainty", "Lat", "Long")
print(xtable(head(sample_data, 10), 
             auto=TRUE,              
             align = c('r', 'l', 'l', 'l', 'r', 'r', 'r', 'r'),
             display = c('d', 's', 's', 's', 'f', 'f', 's', 's')
             ), 
      type="html")
```

--- .class #id 

## Average Temperature Measurement Uncertainty

Temperature **measurement uncertainty** has decreased over time: For example, for New York, United States, the measurements have improved since the early 1900's. Something to consider when choosing a time period to analyze the data.

```{r temperature-uncertainty, echo=FALSE, fig.height = 6, fig.width = 10, fig.align = 'center'}
sample_data <- data0 %>% filter(Country == country & City == 'New York')
sample_data_min_year <- min(sample_data$Year)
sample_data_max_year <- max(sample_data$Year)
plot_title = paste('Average Temperature Uncertainty in', country, 'Between', sample_data_min_year, 'and', sample_data_max_year)    
ggplot(sample_data, aes(x = Year, y = AverageTemperatureUncertainty)) +
    geom_point(shape = 1) +
    geom_smooth() + 
    scale_x_continuous(breaks = seq(floor(sample_data_min_year/10)*10, floor(sample_data_max_year/10)*10 + 10, by = 20)) +
    labs(x = 'Year', y = expression('Average Temperature Uncertainty (' * Degree~degree*C * ')'), title = plot_title) +
    theme(plot.title = element_text(lineheight=.8, face="bold"))            
```

--- .class #id 

## Average Temperature Trend and Hypothesis Testing

Hypothesis H0: The temperature in your city has **NOT** increased.

T.Test result: Rejected for all cities in the United States.

```{r temperature-trend-and-ho-test, echo=FALSE, fig.height = 6, fig.width = 10, fig.align = 'center', warning=FALSE, message=FALSE, error=TRUE} 
country_data <- data0 %>% filter(Country == country)
country_data_grouped_by_city <- country_data %>% group_by(City) %>% summarize(MinYear = min(Year), MaxYear = max(Year))
year_from = 1950
year_to = 2013
min_year = max(country_data_grouped_by_city$MinYear)
max_year = min(country_data_grouped_by_city$MaxYear) 
trendByCountryAndCity(country_data = country_data, country = country, year_from, year_to, min_year, max_year)
```
